---
title: "SwedishMusselAdaptivePotential"
knit: (function(input_file, encoding) {
  out_dir <- '/';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'README.md'))})
author: "Mathias Wegner, AWI"
date: "31/03/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SwedishMusselAdaptivePotential
This markdown file covers the analyses used in Ventura A., Wegner K.M., Dupont S. "Assessing adaptation potential to ocean acidification in blue mussels, Mytilus edulis, from the Swedish west coast" 

```{r functions, include = F}
library(MCMCglmm)
library(tidyverse)
library(gridExtra)

######################################################################################################
##      Global variables
######################################################################################################
iterations.g = 10000000
burnin.g = 1000000
thin.g = 500

######################################################################################################
# MCMCglmm gaussian
# model is a gaussian GLMM with Length as response
# the function will run 3 models to test for differences between cultures and batches:
# model1: deformed ~ pH, random = ~animal + culture + Block
# model2: deformed ~ pH, random = ~animal + culture 
# model3: deformed ~ pH, random = ~animal 


makeMCMC_gaussian_with_pH <- function(thisData) #this function tests for differences in growth between pH environments with random slopes
{
	iterations = iterations.g
	burnin = burnin.g
	thin = thin.g
	
	thisData$ID = thisData$Individual
	
	#standardize Length
	thisData$Length = thisData$Length/sd(thisData$Length)
	
	#make pedigree for animal model
	SIRE =as.character(thisData$Sire)
	lSIRE = length(levels(as.factor(SIRE)))
	DAM =as.character(thisData$Dam)
	lDAM =length(levels(as.factor(DAM)))
	ID =c(levels(as.factor(DAM)),levels(as.factor(SIRE)),as.character(thisData$ID))
	SIRE = c(rep(NA,(lSIRE+lDAM)),SIRE)
	DAM =c(rep(NA,(lSIRE+lDAM)),DAM)

	thisData$animal = thisData$ID
		
	pedigree = cbind(ID,SIRE,DAM)
			
	#make weak, but informative priors for the fixed effect and a fixed matrix for the residual variance, because that needs to be fixed for binomial models
	prior1 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002),G4=list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	prior1_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002),G4=list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	prior1_RS_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002),G2 = list(V = diag(2), nu = 0.002), G3 =list(V = 1, nu = 0.002),G4=list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	
	prior2 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	prior2_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	
	prior3 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	prior4 <- list(G = list(G1 = list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	
	#run MCMCglmm fitting random intercepts for animal, Culture and Block 
	model1=MCMCglmm(Length ~ pH, random = ~animal+Dam+Block+Culture ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior1) 
	model1_RS=MCMCglmm(Length ~ pH, random = ~us(pH):animal+Dam+Block+Culture ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior1_RS) 
	model1_RS_RS=MCMCglmm(Length ~ pH, random = ~us(pH):animal+us(pH):Dam+Block+Culture ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior1_RS_RS) 
	
	
	model2_RS=MCMCglmm(Length ~ pH, random = ~us(pH):animal+Block+Culture ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior2_RS) 	
	model2=MCMCglmm(Length ~ pH, random = ~animal+Block+Culture ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior2) 	
	
	model3=MCMCglmm(Length ~ pH, random = ~animal+Block ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior3) 	
	model4=MCMCglmm(Length ~ pH, random = ~animal ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior4) 	
	
	modelList =list(model1,model1_RS,model1_RS_RS,model2,model2_RS,model3,model4)
	
	return(modelList)
}

#this function tests for differences in growth within pH environments
makeMCMC_gaussian <- function(thisData) 
{
	iterations = iterations.g
	burnin = burnin.g
	thin = thin.g
	
	thisData$ID = thisData$Individual
	
	#standardize Length
	thisData$Length = thisData$Length/sd(thisData$Length)
	
	#make pedigree for animal model
	SIRE =as.character(thisData$Sire)
	lSIRE = length(levels(as.factor(SIRE)))
	DAM =as.character(thisData$Dam)
	lDAM =length(levels(as.factor(DAM)))
	ID =c(levels(as.factor(DAM)),levels(as.factor(SIRE)),as.character(thisData$ID))
	SIRE = c(rep(NA,(lSIRE+lDAM)),SIRE)
	DAM =c(rep(NA,(lSIRE+lDAM)),DAM)

	thisData$animal = thisData$ID
		
	pedigree = cbind(ID,SIRE,DAM)
			
	#make weak, but informative priors for the fixed effect and a fixed matrix for the residual variance, because that needs to be fixed for binomial models
	prior1 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002),G4=list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	prior2 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	prior3 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	prior4 <- list(G = list(G1 = list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002))
	
	#run MCMCglmm fitting random intercepts for animal, Culture and Block 
	model1=MCMCglmm(Length ~ 1, random = ~animal+Dam+Block+Culture ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior1) 
	model2=MCMCglmm(Length ~ 1, random = ~animal+Block+Culture ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior2) 	
	
	model3=MCMCglmm(Length ~ 1, random = ~animal+Block ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior3) 	
	model4=MCMCglmm(Length ~ 1, random = ~animal ,family="gaussian",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior4) 	
	
	modelList =list(model1,model2,model3,model4)
	
	return(modelList)
}

#this function tests for differences in deformation rates between pH environments 
makeMCMC_binomial <- function(thisData) 
{
	iterations = iterations.g
	burnin = burnin.g
	thin = thin.g

	
	thisData$ID = thisData$Individual
	
	#make pedigree for animal model
	SIRE =as.character(thisData$Sire)
	lSIRE = length(levels(as.factor(SIRE)))
	DAM =as.character(thisData$Dam)
	lDAM =length(levels(as.factor(DAM)))
	ID =c(levels(as.factor(DAM)),levels(as.factor(SIRE)),as.character(thisData$ID))
	SIRE = c(rep(NA,(lSIRE+lDAM)),SIRE)
	DAM =c(rep(NA,(lSIRE+lDAM)),DAM)

	thisData$animal = thisData$ID
		
	pedigree = cbind(ID,SIRE,DAM)
			
	#make weak, but informative priors for the fixed effect and a fixed matrix for the residual variance, because that needs to be fixed for binomial models
	prior1 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	prior1_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	
	prior2 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	prior3 <- list(G = list(G1 = list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	
	prior4 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002),G4 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	prior4_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002),G4 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	prior4_RS_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002),G2 = list(V = diag(2), nu = 0.002), G3 =list(V = 1, nu = 0.002),G4 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	
	#run MCMCglmm fitting random intercepts for animal, Culture and Block 
	model1=MCMCglmm(Malformed ~ pH, random = ~animal+Block+Culture ,family="categorical",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior1) 	
	model1_RS = MCMCglmm(Malformed ~ pH, random = ~us(pH):animal+Block+Culture ,family="categorical",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior1_RS) 
	model2=MCMCglmm(Malformed ~ pH, random = ~animal+Block ,family="categorical",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior2) 	
	model3=MCMCglmm(Malformed ~ pH, random = ~animal ,family="categorical",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior3) 	
	model4 = MCMCglmm(Malformed ~ pH, random = ~animal+Dam+Block+Culture ,family="categorical",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior4)
	model4_RS = MCMCglmm(Malformed ~ pH, random = ~us(pH):animal+Dam+Block+Culture ,family="categorical",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior4_RS)
	model4_RS_RS = MCMCglmm(Malformed ~ pH, random = ~us(pH):animal+us(pH):Dam+Block+Culture ,family="categorical",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior4_RS_RS)
	model4_RI_RS = MCMCglmm(Malformed ~ pH, random = ~us(pH):Dam+animal+Block+Culture ,family="categorical",data = thisData,verbose = FALSE,pedigree = pedigree, nit = iterations, burnin = burnin, thin = thin, prior = prior4_RS)

	
	
	modelList =list(model1,model1_RS,model2,model3,model4,model4_RS,model4_RS_RS,model4_RI_RS)
	
	return(modelList)
}


#this function tests for differences in deformation rates between pH environments
makeMCMC_binomial_surv <- function(thisData)  
{
	iterations = iterations.g
	burnin = burnin.g
	thin = thin.g

	
	#make pedigree for animal model
	SIRE =as.character(thisData$Sire)
	lSIRE = length(levels(as.factor(SIRE)))
	DAM =as.character(thisData$Dam)
	lDAM =length(levels(as.factor(DAM)))
	ID =c(levels(as.factor(DAM)),levels(as.factor(SIRE)),as.character(thisData$ID))
	SIRE = c(rep(NA,(lSIRE+lDAM)),SIRE)
	DAM =c(rep(NA,(lSIRE+lDAM)),DAM)

	thisData$animal = thisData$ID
		
	pedigree = cbind(ID,SIRE,DAM)
			
	#make weak, but informative priors for the fixed effect and a fixed matrix for the residual variance, because that needs to be fixed for binomial models
	prior1 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	prior1_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002), G2 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	
	prior2 <- list(G = list(G1 = list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	
	prior3 <- list(G = list(G1 = list(V = 1, nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	prior3_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002),G2 = list(V = 1, nu = 0.002), G3 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	prior3_RS_RS <- list(G = list(G1 = list(V = diag(2), nu = 0.002),G2 = list(V = diag(2), nu = 0.002), G3 =list(V = 1, nu = 0.002)),R = list(V = 1, nu = 0.002, fix = TRUE))
	
	#run MCMCglmm fitting random intercepts for animal, Culture and Block 
	model1=MCMCglmm(cbind(Live,Dead)~ pH, random = ~animal+Block ,family="multinomial2",data = thisData,verbose = FALSE, nit = iterations, burnin = burnin, thin = thin, prior = prior1) 	
	model1_RS = MCMCglmm(cbind(Live,Dead) ~ pH, random = ~us(pH):animal+Block ,family="multinomial2",data = thisData,verbose = FALSE, nit = iterations, burnin = burnin, thin = thin, prior = prior1_RS) 
	model2=MCMCglmm(cbind(Live,Dead) ~ pH, random = ~animal ,family="multinomial2",data = thisData,verbose = FALSE, nit = iterations, burnin = burnin, thin = thin, prior = prior2) 	
	model3 = MCMCglmm(cbind(Live,Dead) ~ pH, random = ~animal+Dam+Block,family="multinomial2",data = thisData,verbose = FALSE, nit = iterations, burnin = burnin, thin = thin, prior = prior3)
	
	model3_RS = MCMCglmm(cbind(Live,Dead) ~ pH, random = ~us(pH):animal+Dam+Block ,family="multinomial2",data = thisData,verbose = FALSE,nit = iterations, burnin = burnin, thin = thin, prior = prior3_RS)
	model3_RS_RS = MCMCglmm(cbind(Live,Dead) ~ pH, random = ~us(pH):animal+us(pH):Dam+Block ,family="multinomial2",data = thisData,verbose = FALSE, nit = iterations, burnin = burnin, thin = thin, prior = prior3_RS_RS)
	
	
	
	modelList =list(model1,model1_RS,model2,model3,model3_RS,model3_RS_RS)
	
	return(modelList)
}

```

## Read Data

```{r read data, include=FALSE}
size <- read.csv("Larvae data for MCMC_2.csv")

surv = read.table("mortality_for_MCMC.txt",header =T)
surv$Family = factor(surv$Family)
surv$Culture = factor(surv$Culture)
surv$Block = factor(surv$Block)
surv$pH = factor(surv$pH)


#convert numerics to factors
size$Individual = factor(size$Individual)
size$Family = factor(size$Family)
size$Culture = factor(size$Culture)
size$Block = factor(size$Block)
size$pH = factor(size$pH)
```

## Larval size
This fit larval size as a function of seawater acidification levels (i.e. pH = 8.1 or pH = 7.5) as fixed effects and “animal” (relatedness between individual larvae), “dam” (female parent), “block” (experimental block) and “culture” (replicate larval culture within each treatment and family) as random intercept effects. To test for family specific reaction norms between the different seawater acidification levels we also fit “animal” as a random slope and did the same for “dam”. We chose the best fitting model based on the respective Deviance Information Criterion (DIC). Only undeformed animals were used.

```{r LengthModels normality}
#remove deformed animals
goodSize=subset(size,size$Malformed == 'no')

#plot distribution with expected normal distribution
p1 = ggplot(data = goodSize, aes(x = Length, group = pH, color = pH, fill = pH)) +
  geom_density(alpha = 0.4) +
  stat_function(fun = function(x) 
    dnorm(x, mean = mean(goodSize$Length[goodSize$pH == 7.5]), sd = sd(goodSize$Length[goodSize$pH == 7.5])) , color = "red", linetype = 2, size = 1.5) +
  stat_function(fun = function(x) 
    dnorm(x, mean = mean(goodSize$Length[goodSize$pH == 8.1]), sd = sd(goodSize$Length[goodSize$pH == 8.1])) , color = "#00BFC4", linetype = 2, size = 1.5) +
  theme_classic() +
  labs(title = "Distribution", y= "Density")

p2 = ggplot(data = goodSize, aes(sample = Length, group = pH, colour = pH, fill = pH)) +
  geom_qq() +
   geom_qq_line() +
  theme_classic() +
  labs(title = "QQ-plot", y= "observed", x = "theoretical quantiles")

grid.arrange(p1,p2, nrow =1)

#test for normality
#within pH 7.5
norm.test_7.5 = ks.test(goodSize$Length[goodSize$pH == 7.5], 'pnorm', 
                        mean = mean(goodSize$Length[goodSize$pH == 7.5]),
                                    sd= sd(goodSize$Length[goodSize$pH == 7.5]))
print(norm.test_7.5)

#within ph 8.1
norm.test_8.1 = ks.test(goodSize$Length[goodSize$pH == 8.1], 'pnorm', 
                        mean = mean(goodSize$Length[goodSize$pH == 8.1]),
                                    sd= sd(goodSize$Length[goodSize$pH == 8.1]))
print(norm.test_8.1)

```

```{r LengthModels runModel, include = F}

LengthModels = makeMCMC_gaussian_with_pH(goodSize)

#give Model Fit
for(i in 1:length(LengthModels))
{
	print('######################################################################################################
')
	print(paste(c('Model:',i)))
	print(summary(LengthModels[[i]]))
}


#calculate models for Length as a response within each pH environment

lowpH = subset(goodSize, goodSize$pH == 7.5)

LengthModels_7.5 = makeMCMC_gaussian(lowpH)

#give Model Fit
for(i in 1:4)
{
	print('######################################################################################################')
	print(paste(c('Model:',i)))
	print(summary(LengthModels_7.5[[i]]))
}


#now for 8.1
highpH = subset(goodSize, goodSize$pH == 8.1)

LengthModels_8.1 = makeMCMC_gaussian(highpH)

#give Model Fit
for(i in 1:4)
{
	print('######################################################################################################')
	print(paste(c('Model:',i)))
	print(summary(LengthModels_8.1[[i]]))
}


best_RS_Model = LengthModels[[5]]
best_75_Model = LengthModels_7.5[[2]]
best_81_Model = LengthModels_8.1[[2]]



autocorr.diag(best_RS_Model$Sol)
autocorr.diag(best_RS_Model$VCV)
#test convergence
heidel.diag(best_RS_Model$VCV)

h2_RS_75 = best_RS_Model$VCV[,'pH7.5:pH7.5.animal']/(best_RS_Model$VCV[,'pH7.5:pH7.5.animal']+best_RS_Model$VCV[,'pH8.1:pH7.5.animal']+best_RS_Model$VCV[,'pH8.1:pH8.1.animal']+best_RS_Model$VCV[,'Block']+best_RS_Model$VCV[,'Culture']+best_RS_Model$VCV[,'units'])
h2_RS_81 = best_RS_Model$VCV[,'pH8.1:pH8.1.animal']/(best_RS_Model$VCV[,'pH7.5:pH7.5.animal']+best_RS_Model$VCV[,'pH8.1:pH7.5.animal']+best_RS_Model$VCV[,'pH8.1:pH8.1.animal']+best_RS_Model$VCV[,'Block']+best_RS_Model$VCV[,'Culture']+best_RS_Model$VCV[,'units'])
h2_75 = best_75_Model$VCV[,'animal']/(best_75_Model$VCV[,'animal']+best_75_Model$VCV[,'Block']+best_75_Model$VCV[,'Culture']+best_75_Model$VCV[,'units'])
h2_81 = best_81_Model$VCV[,'animal']/(best_81_Model$VCV[,'animal']+best_81_Model$VCV[,'Block']+best_81_Model$VCV[,'Culture']+best_81_Model$VCV[,'units'])

posterior.mode(h2_RS_75)
posterior.mode(h2_RS_81)
posterior.mode(h2_75)
posterior.mode(h2_81)

HPDinterval(h2_RS_75)
HPDinterval(h2_RS_81)
HPDinterval(h2_75)
HPDinterval(h2_81)

posterior.mode(best_RS_Model$VCV[,'pH7.5:pH7.5.animal'])
HPDinterval(best_RS_Model$VCV[,'pH7.5:pH7.5.animal'])

posterior.mode(best_RS_Model$VCV[,'pH8.1:pH8.1.animal'])
HPDinterval(best_RS_Model$VCV[,'pH8.1:pH8.1.animal'])

posterior.mode(best_RS_Model$VCV[,'pH7.5:pH8.1.animal'])
HPDinterval(best_RS_Model$VCV[,'pH7.5:pH8.1.animal'])

posterior.mode(best_RS_Model$VCV[,'Culture'])
HPDinterval(best_RS_Model$VCV[,'Culture'])

posterior.mode(best_RS_Model$VCV[,'Block'])
HPDinterval(best_RS_Model$VCV[,'Block'])

posterior.mode(best_RS_Model$VCV[,'units'])
HPDinterval(best_RS_Model$VCV[,'units'])


posterior.mode(best_75_Model$VCV[,'animal'])
HPDinterval(best_75_Model$VCV[,'animal'])

posterior.mode(best_75_Model$VCV[,'Culture'])
HPDinterval(best_75_Model$VCV[,'Culture'])

posterior.mode(best_75_Model$VCV[,'Block'])
HPDinterval(best_75_Model$VCV[,'Block'])

posterior.mode(best_75_Model$VCV[,'units'])
HPDinterval(best_75_Model$VCV[,'units'])

posterior.mode(best_81_Model$VCV[,'animal'])
HPDinterval(best_81_Model$VCV[,'animal'])

posterior.mode(best_81_Model$VCV[,'Culture'])
HPDinterval(best_81_Model$VCV[,'Culture'])

posterior.mode(best_81_Model$VCV[,'Block'])
HPDinterval(best_81_Model$VCV[,'Block'])

posterior.mode(best_81_Model$VCV[,'units'])
HPDinterval(best_81_Model$VCV[,'units'])


plot(best_RS_Model$VCV[,'pH7.5:pH7.5.animal'], main = 'animal RS', ylim = c(0,5))
plot(best_81_Model$VCV[,'animal'], main = 'animal')
```

## Larval Deformation 
This fits a bionmial model analysing deformation as a function of pH, and “animal” (relatedness between individual larvae), “dam” (female parent), “block” (experimental block) and “culture” (replicate larval culture within each treatment and family) as random intercept effects.

```{r deformation }
DeformModels = makeMCMC_binomial(size)
#give Model Fit
for(i in 1:length(DeformModels))
{
	print('######################################################################################################
')
	print(paste(c('Model:',i)))
	print(summary(DeformModels[[i]]))
}

bestDeformedModel = DeformModels[[7]]


h2_def_75 = bestDeformedModel$VCV[,'pH7.5:pH7.5.animal']/(bestDeformedModel$VCV[,'pH7.5:pH7.5.animal']+bestDeformedModel$VCV[,'pH8.1:pH7.5.animal']+bestDeformedModel$VCV[,'pH8.1:pH8.1.animal']+bestDeformedModel$VCV[,'pH7.5:pH7.5.Dam']+bestDeformedModel$VCV[,'pH8.1:pH7.5.Dam']+bestDeformedModel$VCV[,'pH8.1:pH8.1.Dam']+bestDeformedModel$VCV[,'Block']+bestDeformedModel$VCV[,'Culture']+bestDeformedModel$VCV[,'units']+1)
h2_def_81 = bestDeformedModel$VCV[,'pH8.1:pH8.1.animal']/(bestDeformedModel$VCV[,'pH7.5:pH7.5.animal']+bestDeformedModel$VCV[,'pH8.1:pH7.5.animal']+bestDeformedModel$VCV[,'pH8.1:pH8.1.animal']+bestDeformedModel$VCV[,'pH7.5:pH7.5.Dam']+bestDeformedModel$VCV[,'pH8.1:pH7.5.Dam']+bestDeformedModel$VCV[,'pH8.1:pH8.1.Dam']+bestDeformedModel$VCV[,'Block']+bestDeformedModel$VCV[,'Culture']+bestDeformedModel$VCV[,'units']+1)


d2_def_75 = bestDeformedModel$VCV[,'pH7.5:pH7.5.Dam']/(bestDeformedModel$VCV[,'pH7.5:pH7.5.animal']+bestDeformedModel$VCV[,'pH8.1:pH7.5.animal']+bestDeformedModel$VCV[,'pH8.1:pH8.1.animal']+bestDeformedModel$VCV[,'pH7.5:pH7.5.Dam']+bestDeformedModel$VCV[,'pH8.1:pH7.5.Dam']+bestDeformedModel$VCV[,'pH8.1:pH8.1.Dam']+bestDeformedModel$VCV[,'Block']+bestDeformedModel$VCV[,'Culture']+bestDeformedModel$VCV[,'units']+1)
d2_def_81 = bestDeformedModel$VCV[,'pH8.1:pH8.1.Dam']/(bestDeformedModel$VCV[,'pH7.5:pH7.5.animal']+bestDeformedModel$VCV[,'pH8.1:pH7.5.animal']+bestDeformedModel$VCV[,'pH8.1:pH8.1.animal']+bestDeformedModel$VCV[,'pH7.5:pH7.5.Dam']+bestDeformedModel$VCV[,'pH8.1:pH7.5.Dam']+bestDeformedModel$VCV[,'pH8.1:pH8.1.Dam']+bestDeformedModel$VCV[,'Block']+bestDeformedModel$VCV[,'Culture']+bestDeformedModel$VCV[,'units']+1)


posterior.mode(h2_def_75)
posterior.mode(h2_def_81)

HPDinterval(h2_def_75)
HPDinterval(h2_def_81)


posterior.mode(d2_def_75)
posterior.mode(d2_def_81)

HPDinterval(d2_def_75)
HPDinterval(d2_def_81)

posterior.mode(bestDeformedModel$VCV[,'pH7.5:pH7.5.animal'])
HPDinterval(bestDeformedModel$VCV[,'pH7.5:pH7.5.animal'])

posterior.mode(bestDeformedModel$VCV[,'pH8.1:pH8.1.animal'])
HPDinterval(bestDeformedModel$VCV[,'pH8.1:pH8.1.animal'])

posterior.mode(bestDeformedModel$VCV[,'pH7.5:pH8.1.animal'])
HPDinterval(bestDeformedModel$VCV[,'pH7.5:pH8.1.animal'])

posterior.mode(bestDeformedModel$VCV[,'pH7.5:pH7.5.Dam'])
HPDinterval(bestDeformedModel$VCV[,'pH7.5:pH7.5.Dam'])

posterior.mode(bestDeformedModel$VCV[,'pH8.1:pH8.1.Dam'])
HPDinterval(bestDeformedModel$VCV[,'pH8.1:pH8.1.Dam'])

posterior.mode(bestDeformedModel$VCV[,'pH7.5:pH8.1.Dam'])
HPDinterval(bestDeformedModel$VCV[,'pH7.5:pH8.1.Dam'])


posterior.mode(bestDeformedModel$VCV[,'Culture'])
HPDinterval(bestDeformedModel$VCV[,'Culture'])

posterior.mode(bestDeformedModel$VCV[,'Block'])
HPDinterval(bestDeformedModel$VCV[,'Block'])

posterior.mode(bestDeformedModel$VCV[,'units'])
HPDinterval(bestDeformedModel$VCV[,'units'])


```

## Surivival 
This fits a bionmial model analysing deformation as a function of pH, and “animal” (relatedness between individual larvae), “dam” (female parent), “block” (experimental block) and “culture” (replicate larval culture within each treatment and family) as random intercept effects.



```{r survival}
survMCMC = makeMCMC_binomial_surv(surv)
for(i in 1:length(survMCMC))
{
	print('######################################################################################################
')
	print(paste(c('Model:',i)))
	print(summary(survMCMC[[i]]))
}
bestSurvMod = survMCMC[[4]]
plot(bestSurvMod$VCV)

autocorr.diag(bestSurvMod$Sol)
autocorr.diag(bestSurvMod$VCV)
#test convergence
heidel.diag(bestSurvMod$VCV)

h2_surv = bestSurvMod$VCV[,'animal']/(bestSurvMod$VCV[,'animal']+bestSurvMod$VCV[,'Dam']+bestSurvMod$VCV[,'Block']+bestSurvMod$VCV[,'units']+1)
posterior.mode(h2_surv)
HPDinterval(h2_surv)

d2_surv = bestSurvMod$VCV[,'Dam']/(bestSurvMod$VCV[,'animal']+bestSurvMod$VCV[,'Dam']+bestSurvMod$VCV[,'Block']+bestSurvMod$VCV[,'units']+1)
posterior.mode(d2_surv)
HPDinterval(d2_surv)

posterior.mode(bestSurvMod$VCV[,'animal'])
HPDinterval(bestSurvMod$VCV[,'animal'])

posterior.mode(bestSurvMod$VCV[,'Dam'])
HPDinterval(bestSurvMod$VCV[,'Dam'])

posterior.mode(bestSurvMod$VCV[,'Block'])
HPDinterval(bestSurvMod$VCV[,'Block'])




```


